\documentclass[../main.tex]{subfiles}

\begin{document}
In this section we will prove the so called \textit{fixed point theorem} for
\GL.  This theorems says that there for some formulae $\alpha$  of \GL\ exists fixed
points $\sigma$, i.e that for these formuale:
\[\vdash_\GL\sigma\leftrightarrow\alpha(\sigma)\]
This theorem gives is one of the most striking applications of modal logic in
the study of provability in arithemtic; it says that self-reference is not
necessary in the proofs of Gödel's incompleteness theorems.
There has been a lot of different ways to prove this theorem; Boolos
list three different ways in his book \parencite{Boolos1993} and Per Lindström has
another one in his paper \parencite{Lind1996}. Here we
will follow the proof from \parencite{Olson1990}.

This proof of the theorem is a semantical proof; i.e it makes use of the
Kripke semantics of the modal logic \GL, and it further gives a simple
procedure for "calculating" fixed points.

We will end this chapter with calculating some fixed points. This will be
possible since our the proof of the fixed point theorem contains a algorithm,
which can be used to calculate these. In the next chapter we will use Solovay's
completeness theorems and the fixed point theorem to conclude a few things
about arithmetic with the help of modal logic. This is the \textit{power} of
the fixed point theorem; it makes it possible to study the proof predicate with
the help of modal logic.

\section{Prerequisite Results and Definitions}

We will start of by make stating some definitions and results that are crucial
in the proof of the fixed point theorem. We will also in this section define
the special formulae for which the fixed point theorem holds.

\begin{defi}
	We abbreviate $\Box\alpha\wedge\alpha$ as $\sBox\alpha$ for every
	$\alpha$ in our langauge.
\end{defi}
$\sBox$ is called "strong box" and for a given model $\mathcal{K}$ we have that
$\vdash_w^\mathcal{K}\sBox \alpha$ states that $\alpha$ is both true and
provable in node $w$. There are a few theorem about $\sBox$ in the modal logic
\textbf{K4} and logics which extends this, that will be
important in the following. These results are syntactical. We will show these
now, and more results can be found in \parencite{Smor1985} and \cite{Boolos1993}

\begin{thm}
	\label{seet}
	Let $\alpha$ be a modal formula then:
	\begin{enumerate}
		\item $\vdash_{\textbf{K4}}(\sBox\Box\alpha\leftrightarrow\Box\alpha)\leftrightarrow
	\Box\sBox \alpha$
	\item $\vdash_{\textbf{K4}}\sBox\alpha\leftrightarrow\sBox\
		\sBox\alpha$
	\end{enumerate}
\end{thm}
The proof of this is syntactical and is left out.
\begin{thm}
	\label{GLsæt}
	Suppose that \textbf{L} extends \textbf{K4}, and we have that
	$\vdash_{\textbf{L}}\sBox\alpha\rightarrow\beta$. Then
	$\vdash_{\textbf{L}}\Box\alpha\rightarrow\Box\beta$ and
	$\vdash_{\textbf{L}}\sBox\alpha\rightarrow\sBox\beta$
\end{thm}
This theorem clearly holds for \GL, since it is an extension of \textbf{K4} and we will use it later in this chapter.
\begin{proof}
	\textbf{Dobbelt tjek dette}

	Since \textbf{L} extends \textbf{K4} we have that
	$\vdash_{\textbf{L}}\Box\sBox\alpha\rightarrow\Box\beta$ and thus by
	theorem \ref{seet} we have that
	$\vdash_{\textbf{L}}\Box\alpha\rightarrow\Box\beta$ and by the
	definition of $\sBox$ we get that
	$\vdash_{\textbf{L}}\sBox\alpha\rightarrow\sBox\beta$
\end{proof}
Having shown a few results about the syntactical  properties of $\sBox$ we will now
turn to some of the sementical properties about the relationship between $\Box$
and $\sBox$.
\begin{remark}
	\label{rem:acc}
By our semantic of modal logic,  we have that $\vDash_w\sBox\alpha$ is true iff $\vDash_v\alpha$
for all $v\in\{w\}\cup\text{acc}(w)$, where $\text{acc}(w)$ are $\text{acc}(w)$
is the collection of notes that are accessible  from $w$, i.e:
$\text{acc}(w)=\{v\in W:wRv\}$.
\end{remark}

This remark gives us an easy way to determine if $\sBox \alpha$ is true.


Another useful definition that we can make, since we can with loss of
generality issue that  for the  Kripke model 
$\mathcal{K}=\la W,R,\phi\ra$ of \GL\  we look at that the set $W$ is finite,
is the following:
\begin{defi}
	Let $\mathcal{K}=\la W,R,\phi\ra$ be a Kripke model of \GL. The
	\textit{rank} of the nodes $w\in W$ is defined in the following
	way: $\text{rank}(w)=0$ iff there is no world $v$ such that
	$wRv$. Otherwise we have that
	$\text{rank}(w)=1+\text{max}\{\text{rank}(v):wRv\}$.
\end{defi}
So rank$(w)$ simple states how how long the longest path from $w$ is, and the
rank is finite since $W$ is finite.
Further since $W$ is finite and that $R$ is irreflexive we have that the for each $w\in
W$ the rank of $w$ is unique.

We will need the following three lemmas in our proof of the Fixed Point
Theorem.
All three of these lemmas are sementical in nature. 
\begin{lem}
	\label{lem:acc}
	Given any Kripke model $\mathcal{K}=\la W,R,\phi\ra$ of \GL, $w\in W$ and
	sentence $\alpha$, we have that if $\vDash_w\Box\alpha$ then
	$\vDash_v\Box\alpha$ for any $v\in\text{acc}(w)$. Further we have that
	$\vDash_v\sBox\alpha$ for all $v\in\text{acc}(w)$.
\end{lem}
\begin{proof}
	Assume that we have $\vDash_w\Box\alpha$, $w$ sees  $v$ and
	$v$ sees $v'$. Since $R$ is transitive we have that $w$ also sees $v'$ and thus we have that $\vDash_{v'}\alpha$. Since
	$v'$ was chosen arbitraily we have that $\vDash_v\Box\alpha$. We also
	have that $\vDash_v\alpha$ and thus we have $\vDash_v\sBox\alpha$.
\end{proof}
\begin{lem}
	\label{lem:con}
	Given any Kripke model $\mathcal{K}=\la W,R,\phi\ra$ of \GL, $w\in W$ and
	sentence $\alpha$, if $\not\vDash_w\Box\alpha$ then there is "notes"
	$v$ seen by  $w$ such that $\vDash_v\Box\alpha$ and
	$\not\vDash_v\alpha$.
\end{lem}
\begin{proof}
	Assume that $\not\vDash_w\Box\alpha$ then there is a note $v$
	that sees  $w$ such that $\not\vDash_v\alpha$. Let $v$ be the notes
	with the least rank with this property and suppose that $vRv'$. Since
	$v'$ is of less rank than $v$ we have that $\vDash_{v'}\alpha$. Now
	since $v'$ was chosen arbitrarily we have that $\vDash_v\Box\alpha$ and
	the lemma is proven.
\end{proof}
The next lemma will be crucial in the proof of the fixed-point theorem; since
it enable us to substitute two formulae into another formula.
\begin{lem}[Semantic Substitution Lemma]
	\label{lem:sem}
	For any sentences $\alpha,\beta$ and $\gamma$ we have that the
	following formula is valid in all models of \GL:
	$$\sBox(\beta\leftrightarrow\gamma)\rightarrow(\alpha(\beta)\leftrightarrow\alpha(\gamma))$$
\end{lem}
The meaning of $\alpha(\beta)$ is that we replace every occurrences of $p$ in
$\alpha$ with $\beta$; the meaning of $\alpha(\gamma)$ is similar.

\begin{proof}
	We start of by fixing $\beta$ and $\gamma$. The proof will be by
	induction on the complexity of $\alpha$. We will only proof the part
	where $\alpha$ is $\Box\sigma$.

	So suppose that $\alpha$ is $\Box\sigma$, where
	$\sBox(\beta\leftrightarrow\gamma)\rightarrow(\sigma(\beta)\leftrightarrow\sigma(\gamma))$
	is valid. Let $\mathcal{K}$ be any model of \GL\ and let $w\in W$.
	Suppose that $\vDash_w\sBox(\beta\leftrightarrow\gamma)$. Let $v$ be
	any world seen by $w$. Then by lemma \ref{lem:acc} we have
	$\vDash_v\sBox(\beta\leftrightarrow\gamma)$. Since 
	$\sBox(\beta\leftrightarrow\gamma)\rightarrow(\sigma(\beta)\leftrightarrow\sigma(\gamma))$
	is valid we get that
	$\vDash_v\sigma(\beta)\leftrightarrow\sigma(\gamma)$ and since
	$v$ was chosen arbitrary we get
	$\vDash_w\Box(\sigma(\beta)\leftrightarrow\sigma(\gamma))$. By
	proposition \ref{prop:GL} and weak completeness we can conclude
	$\vDash_w\Box\sigma(\beta)\leftrightarrow\sigma(\gamma)$.\textbf{Dobbelt
	tjek om nota er rigtig}.
\end{proof}

We are now almost ready to state state and prove the fixed-point theorem. We
will just need the next two definitions:
\begin{defi}
	A formula $\alpha$ is called \textit{modalized in} $p$ if every occurrence of
	$p$ in  $\alpha$ is under the scope of $\Box$.
\end{defi}
That  $\alpha$ is modalized in $p$ will be a sufficient  condition for $\alpha$
having a fixed point, but it is not a necessary condition for this. $\Box p\vee
p$ is one example of a formula that has a fixed point, even though $p$ is not
modalized in this formula.  This will be showed later.

We will also need the following definition, which in some way is a
generalization of being modalized in $p$:
\begin{defi}
	A sentence $\alpha$ is said to be $n$-\textit{decomposable} iff for some
	sequence $q_1,\ldots,q_n$ consisting of distinct senctence letters that
	do not occur in $\alpha$ we have some sentence $\beta(q_1,\ldots,q_n)$
	that do not contain $p$ and another sequence of distinct sentences 
	$\gamma_1(p),\ldots,\gamma_n(p)$, which each contains $p$ that we have
	$$\alpha=\beta(\gamma_1(p),\ldots,\gamma_n(p))$$.
\end{defi}
The following remark will be important in the proof of the fixed point theorem.
\begin{remark}
It should be noted that if $\alpha$ is modalized in $p$ that we then have that
$\alpha$ is $n$-decomposable for some $n$.
\end{remark}

\section{Proof of the Fixed Point Theorem}

We can now state the fixed point theorem. But we will still need a few lemmas
in our proof. These will be stated and proven when needed. The theorem we will
start of by showing can be seen as a uniqueness theorem, we will after this
theorem state and prove a theorem that can be seen as a existences theorem.
Lastly we will prove a corollary that will show why the name fixed point
theorem is appropriate; this is not exactly from the fixed point theorem
itself.

\begin{thm}
	\label{thm:Fixed}
	If $\alpha$ is modalized in $p$, then there exists a formula $\sigma$
	in which the only sentence letters that occurs are these other than $p$
	that occur in $\alpha$, and such that:
	\[\sBox(p\leftrightarrow\alpha)\rightarrow(p\leftrightarrow\sigma)\]
	The formula $\sigma$ is called a \textit{fixed-point} of $\alpha$.
\end{thm}
We will later on proof that $\vdash_\GL\sigma\leftrightarrow\alpha(\sigma)$ for
such a $\sigma$ and thus the name fixed point is appropriate.
\begin{proof}
	We will prove this by showing that if $\alpha$ is $n$-decomposable
	then it has a fixed point. We will show this by induction on $n$.

	\textbf{Base case:} Suppose that $\alpha$ is $0$-decomposable. Then we
	have that $p$ dose not occur in $\alpha$ and it can thus itself be the
	sentence $\beta$.

	\textbf{Induction step:}
	Assume that every sentence that is $n$-decomposable has a fixed point.
	We now have to show that every sentence that is $(n+1)$-decomposable
	also has a fixed point. To show this we will assume the following:
	$$\alpha(P)=\beta(\Box\gamma_1(p),\ldots,\Box_{n+1}(p))$$
	Further for each $i$ such that $1\leq i\leq n+1$ let:
	\[\alpha_i(p)=\beta(\Box\gamma_1(p),\ldots,\Box\gamma_{i-1},\top,\Box\gamma_{i+1},\ldots,\Box\gamma_{n+1}(p))\]
	Thus we have that for each $i$ such that $1\leq i\leq n+1$ that $\alpha(p)$ is $n$-decomposable,
	so it has a fixed point, that we call $\sigma_i$. Lastly we define:
	\[\sigma=\beta(\Box\gamma_1(\sigma_1),\ldots,\Box\gamma_{n+1}(\sigma_{n+1}))\]
	Our goal is to show that $\sigma$ is a fixed point of $\alpha$, and for
	that we will need the following lemma:
	\begin{lem}
		\label{lem:fix}
		For each $i$ such that $1\leq i\leq n+1$ we have that:
		$$\vdash_\GL\sBox(p\leftrightarrow\alpha)\rightarrow\sBox\Box\gamma_i(p)\leftrightarrow\gamma_i(\sigma_i))$$
	\end{lem}
	\begin{proof}
		Since we have that \GL\ is complete, we just ave to show that
		for any model $\mathcal{K}=\la W,R,\phi\ra$ and any $w\in W$
		that:
		\begin{equation}
		\label{eqn:1}
		\mathcal{K}\vDash\sBox(p\leftrightarrow\alpha)\rightarrow(\sBox(\Box\gamma_i(p)\leftrightarrow
		\Box\gamma_i(\sigma_i))
	\end{equation}
		So we will start of by fixing $i$, $\mathcal{K}$ and $w\in W$.
		We will show \ref{eqn:1} by assuming
		$\vDash_w\sBox(p\leftrightarrow\alpha)$ and then deduce:
		$\vDash_w\sBox(\Box\gamma_i(p)\leftrightarrow\Box\gamma_i(\sigma_i))$;
		this is equivalent to $\vDash_v\Box
		\gamma_i(p)\leftrightarrow\Box\gamma_i(\sigma_i)$ for all
		$v\in\{w\}\cup\text{acc}(w)$ by remark \ref{rem:acc}. So let
		$v\in\{w\}\cup\text{acc}(w)$ and assume that
		$\vDash\Box\gamma_i(p)$, i.e
		$\vDash_v\Box\gamma_i(p)\leftrightarrow\top$. By lemma
		\ref{lem:acc} we have that for any $v'\in\text{acc}(v)$ that
		$\vDash_{v'}\Box\gamma_i(p)$ and thus
		$\vDash_{v'}(\Box\gamma_i(p)\leftrightarrow\top$). This means
		that we have:
		\[\vDash_v\sBox(\Box\gamma_i(p)\leftrightarrow\top)\]
		And thus by lemma \ref{lem:sem} we get that
		$\vDash_v\alpha_i\leftrightarrow\alpha$ and since our $v$ was
		chosen arbitrarily we have that
		$\vDash_w\Box(\alpha_i\leftrightarrow\alpha)$ and thus by
		lemma \ref{lem:acc} we get that:
		$\vDash_v\sBox(\alpha_i\leftrightarrow\alpha)$. Since we have
		assumed that $\vDash_w\sBox(p\leftrightarrow\alpha)$ we again
		have by lemma \ref{lem:acc} that
		$\vDash_v\sBox(p\leftrightarrow\alpha)$, and hence we have
		$\vDash_v\sBox(p\leftrightarrow\alpha_i)$. Since our logic is
		complete and we have assumed by the induction hypothesis that
		$\alpha_i$ has a fixed point $\sigma_i$ we have that
		$\vDash_v(p\leftrightarrow\gamma_i)$, and thus, since $v$ was
		chosen arbitrarily we have that
		$\vDash_w\Box(p\leftrightarrow\gamma_i)$. So by using
		\ref{lem:acc} again we get that
		$\vDash_v\sBox(p\leftrightarrow\gamma_i)$. We will now use
		lemma \ref{lem:sem} again and get that:
		\begin{equation}
			\label{eqn:sub1}
			\vDash_v\gamma_i(p)\leftrightarrow\gamma_i(\sigma_i)
		\end{equation}
		and
		\begin{equation}
			\label{eqn:sub2}
			\vDash_v\Box\gamma_i(p)\leftrightarrow\Box\gamma_i(\sigma_i)
		\end{equation}
		Notice that these two holds for any $v\in\{w\}\cup\text{acc}(w)$ such that
		$\vDash_v\Box\gamma_i(p)$.
		Further by \ref{eqn:sub2} we can deduce
		$\vDash_v\Box\gamma_i(p)\rightarrow\Box\gamma_i(\sigma_i)$

		For the next step of the prove of this lemma we will assume that
		$\not\vDash_v\Box\gamma_i(p)$. This means by lemma
		\ref{lem:con} that there is some world $v'$ where
		$v'\in\text{acc}(v)$, such that $\not\vDash_{v'}\gamma_i(p)$
		and $\vDash_{v'}\Box\gamma_i(p)$. \ref{eqn:sub1} holds for $v'$
		since $v'\in\{w\}\cup\text{acc}(w)$ and thus we have
		$\vDash_{v'}\gamma_i(p)\leftrightarrow\gamma_i(\sigma_i)$. This
		gives that $\not\vDash_{v'}\gamma_i(\sigma_i)$ and thus since
		$vRv'$ we have that $\not\vDash_v\Box\gamma_i(\sigma_i)$. By
		contraposition we then get:
		$\vDash_v\Box\gamma_i(\sigma_i)\rightarrow\Box\gamma_i(p)$, and
		thus we have shown that:
		$$\vDash_v\Box\gamma_i(p)\leftrightarrow\Box\gamma_i(\sigma_i)$$
		We have now shown the lemma.
	\end{proof}
	We now go back and finish our proof of the fixed point theorem. Suppose
	that $\mathcal{K}$ is a model and that $w\in W$ such that
	$\vDash_w\sBox(p\leftrightarrow\alpha)$. By lemma \ref{lem:fix} and
	completeness we get
	$\vDash_w\sBox(\Box\gamma_i(p)\leftrightarrow\Box\gamma_i(\sigma_i))$.
	By using lemma \ref{lem:sem} $(n+1)$ times we can deduce that:
	$$\vDash_w\beta(\Box\gamma_i(p),\ldots,\Box\gamma_{n+1}(p))\leftrightarrow\beta(
	\Box\gamma_1(\sigma_1),\ldots,\Box\gamma_{n+1}(\sigma_{n+1}))$$
	i.e $\vDash_w\alpha\leftrightarrow\sigma$. 

	Since we have $\vDash_w p\leftrightarrow\alpha$ we get $\vDash_w
	p\leftrightarrow\sigma$, we can obtain
	$\vDash_w\sBox(p\leftrightarrow\alpha)\rightarrow(p\leftrightarrow\sigma)$.
	Since our $\mathcal{K}$ and $w$ was chosen arbitrarily we have that
	$\sBox(p\leftrightarrow\alpha)\rightarrow(p\leftrightarrow\sigma)$ is
	valid. By completeness we then have:
	$\vdash_\GL\sBox(p\leftrightarrow\alpha)\rightarrow(p\leftrightarrow
	\sigma)$
\end{proof}

The following result follows from the fixed-point theorem.
\begin{thm}\label{thm:exi}
	Let $\alpha(p)$ be modalized in $p$, and let $\sigma$ be a fixed-point
	of $\alpha$. Then:
	\[\vdash_\GL\sBox(p\leftrightarrow\sigma)\rightarrow(p\leftrightarrow\alpha)\]
\end{thm}
\begin{proof}
	Suppose that $\mathcal{K}=\la W,R,\phi\ra$ is a finite transitive and irreflexive model
	in which
	$\sBox(p\leftrightarrow\sigma)\rightarrow(p\leftrightarrow\alpha)$ is
	invalid. This means that for some $w\in W$ of least rank that we have:
	$\vDash_w\sBox(p\leftrightarrow\sigma)$ and thus $\vDash_w
	p\leftrightarrow\sigma$ and $\not\vDash_w p\leftrightarrow\alpha$. If $wRv$
	then $\vDash_v\sBox(p\leftrightarrow\sigma)$ and since $x$ of lower
	rank than $w$ we also have $\vDash_v p\leftrightarrow\alpha$. Let
	$\varphi'$ be like $\varphi$ expect that $p\in\phi'(w)$ if and only if
	not $p\in\phi(w)$. Set $\mathcal{K}'=\la W,R,\phi'\ra$, and this model
	is clearly transitive and irreflexive.

	In the rest of this prove we will use corollary to the continuity
	theorem; i.e corollary \ref{cor:conti}.


	The formula $\alpha$ is a truth-functional compound  of closed formulas $\Box\beta$ and proportional letters $q$
	such that each $q$ is not $p$. We have that
	$\vDash_w^\mathcal{K}\Box\beta$ if and only if
	$\vDash_v^\mathcal{K}\beta$ for all $v$ such that $wRv$, if and only if
	$\vDash_v^{\mathcal{K}'}$ for all $v$ such that $wRv$ (by continuity),
	if and only if $\vDash_w^{\mathcal{K}'}\Box\beta$. We further have by
	the definition of $\mathcal{K}'$ that $\vDash_w^\mathcal{K}\alpha$ if
	and only iff $\vDash_w^{\mathcal{K}'}$ and $\vDash_w^\mathcal{K}p$ if
	and only if not $\vDash_w^{\mathcal{K}'}p$. Thus we have
	$\vDash_w^{\mathcal{K}'}p\leftrightarrow\alpha$ and b the continuity  theorem
	we again get $\vDash_v^{\mathcal{K}'}p\leftrightarrow\alpha$ for all
	$v$ such that $wRv$. But this means that we get
	$\vDash_w^{\mathcal{K}'}\sBox(p\leftrightarrow\alpha)$.
	
	Since $\sigma$ does not contain $p$ we get by the conti them that
	$\vDash_w^\mathcal{K}\sigma$ if and only if
	$\vDash_w^{\mathcal{K}'}\sigma$. Since we have that
	$\vDash_w^\mathcal{K} p$ if and only if not $\vDash_w^{\mathcal{K}'}p$
	we get $\not\vDash_w^{\mathcal{K}'}p\leftrightarrow \sigma$ and thus
	$\sBox(p\leftrightarrow\sigma)\rightarrow(p\leftrightarrow\sigma)$ is
	invalid.

	By soundness and completeness of \GL\ we thus get that if
	$\vdash_\GL\sBox(p\leftrightarrow\alpha)\rightarrow(p\leftrightarrow\sigma)$
	then
	$\vdash_\GL\sBox(p\leftrightarrow\sigma)\rightarrow(p\leftrightarrow\alpha)$.
\end{proof}
From this we can prove the following corollary that show that the name
fixed-point is appropriate:
\begin{cor}
	Let $\alpha(p)$ be modalized in $p$ and let $\sigma$ be a fixed-point
	of $\alpha$. Then:
	\[\vdash_\GL\sigma\leftrightarrow\alpha(\sigma)\]
\end{cor}
\begin{proof}
	Since we have assumed uniform substitution and by theorem \ref{thm:exi}
	the result of substituting $\sigma$ for $p$ in
	$\sBox(p\leftrightarrow\sigma)\rightarrow(p\leftrightarrow\alpha)$ is a
	theorem of \GL. This means that
	$\vdash_\GL\sBox(\sigma\leftrightarrow\sigma)\rightarrow(\sigma\leftrightarrow\alpha(\sigma))$.
	$\sBox(\sigma\leftrightarrow\sigma)$ is obviously a theorem of \GL\ so
	we get:
	\[\vdash_\GL\sigma\leftrightarrow\alpha(\sigma)\]
\end{proof}
We can also state  (by theorem \ref{thm:exi} and \ref{thm:Fixed})the fixed
point theorem in the following form:

\begin{cor}
	\label{cor:Fixed}
	For every modal formula $\alpha$ modalized in $p$, there is a modal formula
	$\sigma$ only contaning propostional letters contained in $\alpha$ not
	contaning $p$ such that:
	\[\vdash_\GL\sBox(p\leftrightarrow\alpha)\leftrightarrow\sBox(p\leftrightarrow\sigma)\]
\end{cor}
\section{Calculation of Fixed Points}

The algorithm in the proof of the fixed theorem can be used to calculate
specific fixed points for some given formula, that fulfills the conditions of
the theorem. 

When we have shown Solovay's completeness theorems, the fixed point theorem
will then make it possible for us to determine the fixed point of a wide range
of sentences of arithmetic. This is important because this way of finding these
points are often a lot easier than finding the points only with the help of the
arithmetical methods.

Below are a few example of how we can use this algorithm.

\begin{exmp}
	Let $\alpha(p)=\Box \neg p$. We will calculate the fixed point for
	$\alpha$. Let $\gamma_1 (p)=p$ and $\beta(q_1)=q_1$. We then have that
	$\alpha(p)=\beta(\Box\gamma_1(p))$ and $\alpha_1(p)=\beta(\top)=\top$.
	The fixed point of $\sigma_1$ of $\alpha_1$ is then $\top$, since $p$
	does not occur in $\alpha_1$. We can thus conclude that the fixed point
	$\sigma$ of $\alpha$ is the following:
	\[\beta(\Box\gamma_1(\top))=\beta(\Box\neg\top)=\Box\neg\top\]
	this formula is equivalent to $\Box\bot$.
\end{exmp}
\begin{exmp}
	Let $\alpha(p)=\Box p\rightarrow\Box\neg p$. To find the fixed point,
	let$\gamma_1(p)=p$,$\gamma_2(p)=\neg p$ and
	$\beta(q_1,q_2)=q_1\rightarrow q_2$. Then clearly we have:
	\[\alpha(p)=\beta(\Box\gamma_1(p),\Box\gamma_2(p))\]
	\[\alpha_1(p)=\beta(\top,\Box\gamma_2(p))=\top\rightarrow\Box\neg p\]
	and 
	\[\alpha_2(p)=\beta(\Box\gamma_1(p),\top)=\Box p\rightarrow \top\]
	It is clear that $\alpha_1$ is equivalent to $\Box\neg p$ and that
	$\alpha_2$ is equivalent top $\top$; thus they have fixed points
	$\sigma_1=\Box\bot$ and $\sigma_2=\top$. So the fixed point $\sigma$
	of $\alpha$ is the following:
	\[\beta(\Box\gamma_1(\Box\bot),\Box\gamma_2(\top))=\Box\Box\bot\rightarrow\Box\neg\top\]
	or equivalent: $\Box\Box\bot\rightarrow\Box\bot$.
\end{exmp}
\subsection{Calculation of a Non-Modalized Fixed Point}
We will now show that $\Box p\vee p$ has the fixed point from the table. It is
clear that this formula is not modalized in $p$, but it can still be shown to
have a fixed point.

For
this end we will need the following theorem:
We will start of by showing that no $\beta$ modalized in $p$ is equivalent to
$\Box p\vee p$. This will the following two propositions show:
\begin{prop}
	Suppose that $\beta$ is modalized in $p$. Then we have:
	\[\vdash_\GL p\rightarrow\beta\Rightarrow\ \vdash_\GL\beta\]
\end{prop}
\begin{proof}
	Suppose that $\not\vdash_\GL\beta$. This means that for some finite
	transitive and irreflexive model $\mathcal{K}=\la W,R,\phi\ra$ and some $w\in W$ that:
	$\not\vDash_w\beta$. Let $\phi'$ be such that $w\in \phi'(p)$ and
	otherwise it is just $\phi$ and let $\mathcal{K}'=\la W,R,\phi'\ra$. We
	have that $\beta$ is a truth-functional compound of formulae
	$\Box\gamma$ and proportional letters that are not $p$. By continuity
	we have that $\not\vDash_w^{\mathcal{K}'}\beta$ but we have that
	$\vDash_w^{\mathcal{K}'}p$. Hence we have that
	$\vDash_w^{\mathcal{K}'}p\rightarrow\beta$ and the propositions follows.
\end{proof}
The next proposition shows that no $\beta$ equivalent to $\Box p\vee p$ is
modalized in $p$.
\begin{prop}
	For no $\beta$ modalized in $p$, we have that: 
	\[\vdash_\GL\beta\leftrightarrow(\Box p\vee p)\]
\end{prop}
\begin{proof}
If there was such a $\beta$ we would have that $\vdash_\GL p\rightarrow\beta$
and thus from the just shown proposition we would have that $\vdash_\GL\beta$
and $\vdash_\GL\Box p\vee p$. By substitution we would then have that
$\vdash_\GL\Box\bot\vee\bot$ and this is clearly false.
\end{proof}
So $\Box p\vee p$ is not equivalent to any formula modalized in $p$, but we can
show that it has a fixed point any way:
\begin{prop}
	\[\vdash_\GL\sBox(p\leftrightarrow \Box p\vee
	p)\leftrightarrow\sBox(p\leftrightarrow\top)\]
\end{prop}
\begin{proof}
	The right to left implication is trivial. The other way will be shown
	by the following deduction:
	\begin{align*}
		\vdash_\GL\sBox(p\leftrightarrow\Box p\vee
		p)&\rightarrow\sBox(\Box p\rightarrow p)\\
		  &\rightarrow\sBox\sBox(\Box p\rightarrow p)\ \text{By
		  theorem \ref{seet}}\\
		  &\rightarrow\sBox (\Box(\Box p\rightarrow p)\wedge(\Box
		  p\rightarrow p))\ \text{By definition of}\ \sBox\\
		  &\rightarrow\sBox(\Box p\wedge(\Box p\rightarrow p))\\
		  &\rightarrow\sBox p\\
		  &\rightarrow (p\leftrightarrow \top)
	\end{align*}
\end{proof}
In figure \ref{Fig} we can see some more fixed points. 
formula on the left is the formula $\alpha(p)$ and the formula on the right is
the formula $\sigma$. Line (3) and (6) has already been shown.
\newpage
\begin{align*}
	1. &\ \neg\Box p & \neg\Box\bot\\
	2. &\ \Box p & \top\\
	3. &\ \Box\neg p & \Box\bot\\
	4. &\ \neg\Box\neg p & \neg\Box\bot\\
	5. &\ \neg\Box\Box\neg p & \neg\Box\Box\bot\\
	6. &\ \Box p \rightarrow\Box\neg p & \Box\Box\bot\rightarrow\Box\bot\\
	7. &\ \Box (\neg
	p\rightarrow\Box\bot)\rightarrow\Box(p\rightarrow\Box\bot) &
	\Box\Box\Box\bot\rightarrow\Box\Box\bot\\
	8. &\ \Box p\rightarrow q & \Box q\rightarrow q\\
	9. &\ \Box(p\rightarrow q) & \Box q\\
	10. &\ \Box p\wedge q & \Box q\wedge q\\
	11. &\ \Box(p\wedge q) & \Box q\wedge q\\
	12. &\ q\vee \Box p & \top\\	
	13. &\ \neg\Box(q\rightarrow p) & \Diamond q\\
	14. &\ \Box(p\rightarrow q)\rightarrow\Box\neg p &
	\Box(\Box\bot\rightarrow q)\rightarrow\Box\bot\\
	15. &\ q\wedge(\Box(p\rightarrow q)\rightarrow\Box\neg p) &
	q\wedge\Box\neg q\\
	16. &\ \Diamond p\rightarrow(q\wedge\neg\Box(p\rightarrow q)) &
	\Diamond\top \rightarrow (q\wedge\neg\Box(\Box\bot\rightarrow q))\\
	17. &\ \Box(\Box(p\wedge q)\wedge\Box(p\wedge r)) & \Box(\Box q \wedge
	\Box r)\\
	18. &\ \Box p\vee p & \top
\end{align*}
\begingroup\vspace*{-\baselineskip}
\captionof{figure}{A Table of Fixed Points}
\label{Fig}
\vspace*{\baselineskip}\endgroup

We will after the proof of the soundness part of  Solovay's Completeness Theorems come back to this
table, and make some conclusions about the fixed points.
\end{document}
